\chapter{Introduction}
\section{Problem Statement}

This project involves the design and implementation of an 8051-compatible microcontroller 
using Verilog HDL on a ZyboZ7 FPGA development board. The 8051 architecture is well known by
the class and serves as a practical foundation for applying digital design concepts. 
The project is distributed across the entire class, with themes rotating every two weeks. 
In addition to the core components—peripherals, single-cycle CPU, and pipelined CPU—a final 
theme will involve the development of a simple C compiler targeting the designed 
microcontroller.

The project is divided into two main versions of the microcontroller architecture:
\begin{enumerate}
    \item \textbf{Single-Cycle Version}
    \item \textbf{Pipelined Version}
\end{enumerate}

Each version must be capable of executing a defined subset of the 8051 instruction set and 
interfacing with a set of peripherals also developed by the class. The final implementation
 should demonstrate correct execution of programs and proper interaction with these
  peripherals through memory-mapped I/O or equivalent interfacing strategies.

Both architectures must include:
\begin{itemize}
    \item Instruction fetch, decode, execution, and write-back stages (as applicable)
    \item A register file, ALU, and control logic
    \item Memory interface for instruction and data access
    \item Access to a shared peripheral interface/bus
\end{itemize}

This project will be completed in phases, with key milestones to be defined by the
 Professor and class schedule. Documentation, simulation, synthesis, and on-board 
 testing are required deliverables.

\vspace{0.5cm}
The C-compiler for this 8051 will be presented in a separate documentation.
\vspace{0.5cm}

This project follows a \textbf{waterfall development approach}, 
where each stage of the system's creation—analysis, design, implementation, testing, 
and deployment—is.addressed in sequence. The purpose of this chapter is to analyze the functional 
and structural requirements of the 8051-compatible microcontroller and its subsystems, which 
include the single-cycle CPU, pipelined CPU, peripheral interfaces, and C compiler backend.


\section {Requirements}

\section {Constraints}

\section{System Overview}